Relationships in the extremes and their influence on competition and coexistence: Supporting information

### 2021-09-03 paper/Paper.Rnw
edits

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=2cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{xr}
\usepackage{relsize}
\usepackage{tikz}

@@

We here pursue goal G1 from the Introduction, to augment the MTC to show how ATA
plays a role.

For pedgogical clarity, we develop each step of the theory in general and for a simple example model, the lottery model.

The lottery model is as follows. 
%***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year.
For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
properties, corresponding to various tail association cases. Note that $N$ is constant through time.

Modern coexistence theory \citep{Chesson_2000} and its recent computational extensions \citep{Ellner_2016,Ellner_2019}

quantify the contributions of 
different mechanisms to species coexistence. But we show below
that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
We now define storage effects for a general two species model, and specifically 
for the lottery model, using an approach and notation
similar to those of \cite{Ellner_2016;Ellner_2019}; see those papers for additional details.

@@

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig3.pdf}
\caption{The contribution of ATAs to species coexistence is meaningfully large compared to 
other components of the MTC decomposition. All panels show the IGR and the decomposition of
the IGR provided by the extended MTC, equation (\ref{eq:full_decomp}). For each panel, 
$\delta$ and $\mu_1-\mu_2$ 
are fixed at values specified in the margins, and the IGR and components of the decomposition
are plotted against $\sigma$. The contribution of ATAs (red line) frequently 
has magnitude comparable to other components, indicating that ATAs can contribute meaningfully
to species coexistence or failure to coexist. In some instances, the IGR line (orange) falls 
betweed the ATA line (red)
and the $x$ axis, indicating that, for those parameter values, ATAs make the difference between species
coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities 
were never greater than \Sexpr{round(se3,5)}, so are not displayed. \label{fig:fig3}}
\end{figure}
%***DAN: Dan has edited this caption

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig4.pdf}
\caption{ATAs can sometimes make the difference between species coexistence and exclusion of the weaker
competitor; both ``ATA rescue'' and ``ATA exclusion'' (see Methods) are possible. 
All panels show the IGR, the ATA term of the extended MCT [$\Delta_i^{[E || C]}$ in equation 
(\ref{eq:full_decomp})], and the difference of these quantities. 
For each panel, $\delta$ and $\sigma$ are fixed at values specified in the 
margins, and the three quantities are plotted against $\mu_1-\mu_2$ to isolate points in the
``trait-difference space'' corresponding to ATA rescue 
($\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$) 
or ATA exclusion ($\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$; 
see Methods). Values of $\mu_1-\mu_2$ for which the IGR is positive but 
$\text{IGR}-\Delta_i^{[E || C]}$ is negative correspond to ATA rescue (shaded yellow), whereas
values of $\mu_1-\mu_2$ for which the IGR is negative but $\text{IGR}-\Delta_i^{[E || C]}$ is positive 
correspond to ATA exclusion (shaded violet). Top panels also include domains
of ATA rescue or exclusion, but they were too narrow to be highlighted by shading and are
probably of minimal practical importance. 
Standard errors of plotted quantities were never greater than \Sexpr{round(se4,5)}, 
so are not displayed. The dashed red line on panel L corresponds to the parameters
used for Fig. \ref{fig:pedagog}. This plot uses $q_{ij}=1$. An analogous plot was generated 
using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$, but it looked exactly the same, by mathematical necessity (see SI section ).}\label{fig:fig4}
\end{figure}
%***DAN: Dan has edited this caption

>>>>

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=1.5cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{xr}
\usepackage{relsize}
\usepackage{tikz}

@@

We here pursue goal G1 from the Introduction, to augment the MTC to show how ATA
plays a role. For pedgogical clarity, we develop each step of the theory in general and for 
a simple example model, the lottery model. 
The lottery model is as follows. 
%***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year.
For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
properties, corresponding to various tail association cases. Note that $N$ is constant through time.

Modern coexistence theory \citep{Chesson_2000} and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
quantify the contributions of 
different mechanisms to species coexistence. But we show below
that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
We now define storage effects for a general two species model, and specifically 
for the lottery model, using an approach and notation
similar to those of \cite{Ellner_2016,Ellner_2019}; see those papers for additional details.

@@

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig3.pdf}
\caption{The contribution of ATAs to species coexistence is often meaningfully large compared to 
other components of the MTC decomposition. All panels show the IGR and the decomposition of
the IGR provided by the extended MTC, equation (\ref{eq:full_decomp}). For each panel, 
$\delta$ and $\mu_1-\mu_2$ 
are fixed at values specified in the margins, and the IGR and components of the decomposition
are plotted against $\sigma$. The contribution of ATAs (red line) frequently 
has magnitude comparable to other components, indicating that ATAs can contribute meaningfully
to species coexistence or failure to coexist. In some instances, the IGR line (orange) falls 
betweed the ATA line (red)
and the $x$ axis, indicating that, for those parameter values, ATAs make the difference between species
coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities 
were never greater than \Sexpr{round(se3,5)}, so are not displayed. This plot uses $q_{ij}=1$. An 
analogous plot (SI Fig. \ref{SI-fig:fig3_alt}) was generated 
using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$. Though results differed
in detail, main conclusions were the same. \label{fig:fig3}}
\end{figure}
%***DAN: Dan has edited this caption

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig4.pdf}
\caption{ATAs can sometimes make the difference between species coexistence and exclusion of the weaker
competitor; both ``ATA rescue'' and ``ATA exclusion'' (see Methods) are possible. 
All panels show the IGR, the ATA term of the extended MCT [$\Delta_i^{[E || C]}$ in equation 
(\ref{eq:full_decomp})], and the difference of these quantities. 
For each panel, $\delta$ and $\sigma$ are fixed at values specified in the 
margins, and the three quantities are plotted against $\mu_1-\mu_2$ to isolate points in the
``trait-difference space'' corresponding to ATA rescue 
($\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$) 
or ATA exclusion ($\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$; 
see Methods). Values of $\mu_1-\mu_2$ for which the IGR is positive but 
$\text{IGR}-\Delta_i^{[E || C]}$ is negative correspond to ATA rescue (shaded yellow), whereas
values of $\mu_1-\mu_2$ for which the IGR is negative but $\text{IGR}-\Delta_i^{[E || C]}$ is positive 
correspond to ATA exclusion (shaded violet). Top panels also include domains
of ATA rescue or exclusion, but they were too narrow to be highlighted by shading and are
probably of minimal practical importance. 
Standard errors of plotted quantities were never greater than \Sexpr{round(se4,5)}, 
so are not displayed. The dashed red line on panel L corresponds to the parameters
used for Fig. \ref{fig:pedagog}. This plot uses $q_{ij}=1$. An analogous plot was generated 
using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$, but it looked exactly the same, by mathematical necessity (see SI section \ref{SI-sect:F4andqij}).}\label{fig:fig4}
\end{figure}
%***DAN: Dan has edited this caption

### 2021-09-03 paper/SupMat.Rnw
edits

\begin{document}
\SweaveOpts{concordance=TRUE}
\SweaveOpts{concordance=TRUE}

@@

@@

>>>>

\begin{document}
\SweaveOpts{concordance=TRUE}

@@

\tableofcontents
\listoffigures

@@

\section{Fig. \ref{MT-fig:fig4} does not depend on the value of $q_{ij}$ used}\label{sect:F4andqij}

%***DAN: Jasmin mentioned she had some math proving this, it needs to be added here.

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig3_qij.pdf}
\caption[Analogous to Fig. \ref{MT-fig:fig3} but making a different choice for $q_{ij}$]
{Analogous to Fig. \ref{MT-fig:fig3} of the main text, but using 
$q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$. See the caption of
that figure for details. \label{fig:fig3_alt}}
\end{figure}
%***DAN: Jasmin, in the caption of Fig. 3 in the main text there is a statment about standard
%error. We need another such statement here, but for the standard errors that applied here.
